\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-13 doc/paper.tex
misc last minute updates

	\section{Introduction} Spreadsheets surround us. They are vital to
	industry and academia alike, supporting as much as 95\% of U.S. financial firms
	use them daily~\cite{panko2008sarbanes} and empowering teachers to track
	research and students' grades~\cite{ko2011state}. This flexibility boosts their
	allure: while a novice can work with basic offerings, advanced users can choose
	among the hundreds of specialized functions to fulfill a wide range of needs,
	from complex statistics to text manipulation~\cite{nardi1990spreadsheet}. As
	such, it should be no surprise that tens of millions of workers rely on these
	programs daily~\cite{scaffidi2005estimating}.

	However, this ubiquity is not without danger; as more people make larger and
	more complex spreadsheets, the cost of failure grows too. The European
	Spreadsheet Risks Interest Group (EuSpRIG) documents many horror stories of
	spreadsheets gone wrong\footnote{http://www.eusprig.org/horror-stories.htm},
	where consequences range from the profuse loss of money to the inaccurate
	outcomes of elections to the creation of extensive political initiatives based
	on flawed economics. Even if most spreadsheet errors are relatively benign,
	poor spreadsheet practice can risk millions, if not billions, of
	dollars~\cite{powell2009impact}.

	Fortunately, spreadsheet researchers address these dangers by amassing
	spreadsheet collections from diverse sources, industrial to academic, so
	they can study how people use spreadsheets. Collections like
	EUSES~\cite{fisher2005euses}, Fuse~\cite{barik2015fuse}, and the Enron
	corpus~\cite{hermans2015enron} have already produced valueable research, such
	as the documentation of code smells in spreadsheet
	environments~\cite{hermans2012detecting}~\cite{jansen2015code} or the creation
	of grammars to describe every possible spreadsheet program in
	Excel~\cite{aivaloglou2015grammar}. Additionally, some research compares these
	collections to discover variation in spreadsheet techniques over time and
	location~\cite{jansen2015enron}.

	The value of spreadsheets demands that we design tools to better understand
	these datasets and simplify complexities we find in them. One such complexity
	is formula construction; functions are an important part of using spreadsheets
	efficiently, yet the number of potential combinations is massive considering
	that Excel boasts over 400 functions. Tools that explore these combinations in
	practice can aid researchers in understanding how people program and in
	educating new generations of spreadsheets users on how to work best.

	This paper contributes such a tool, \toolnameend, that attempts to fulfill this
	need for better exploration of spreadsheet corpora. Working with spreadsheets
	in Excel, \toolname focuses primarily on the functions, like SUM and IF and
	hundreds more, which are the building blocks of formulae that determine values
	across a spreadsheet. The tool creates an interactive visualization which
	captures how people combine these functions, quantifying the broad patterns of
	function nesting while also capturing anomalous formulae and their locations.

@@

	Other tools focus on creating visualizations external to the sheets. Hermans
	and colleagues address a spreadsheet programmer's information needs by making
	dataflow diagrams from individual sheets through the tools
	GyroSAT~\cite{hermans2011supporting} and Breviz~\cite{hermans2011breviz}.
	\toolname is among this type, creating standalone visualizations about
	spreadsheets instead of drawing connections inside the spreadsheet itself. In
	contrast to the other visualization work, our work focuses on representing
	functions and how they combine into complex formulae, rather than exploring the
	data structures of any particular spreadsheet.

@@

	\subsection{Design Goals} \label{goals} In visualizing the spreadsheet data, we
	set out a few core goals for what the tool should accomplish. Furthermore, with
	these goals, we also considered the likely challenges of achieving those goals
	and particular outcomes which we considered beyond the scope of the research.
	We present both sides of these goals below:

>>>>

	\section{Introduction} Spreadsheets surround us. They are vital to
	industry and academia alike, supporting as much as 95\% of U.S. financial firms
	daily~\cite{panko2008sarbanes} and empowering teachers to track research and
	students' grades~\cite{ko2011state}, among many other tasks. This flexibility
	grants them their allure: while a novice can work with basic offerings, advanced
	users can choose among the hundreds of specialized functions to fulfill a wide
	range of needs, from complex statistics to text
	manipulation~\cite{nardi1990spreadsheet}. As such, it should be no surprise
	that tens of millions of workers rely on these
	programs daily~\cite{scaffidi2005estimating}.

	However, this ubiquity is not without danger. As more people make larger and
	more complex spreadsheets, the cost of failure grows too. The European
	Spreadsheet Risks Interest Group (EuSpRIG) documents many horror stories of
	spreadsheets gone
	wrong\footnote{\href{http://www.eusprig.org/horror-stories.htm}{http://www.eusprig.org/horror-stories.htm}}, where consequences range from the profuse loss of money to the inaccurate outcomes of elections to the creation of extensive political initiatives based on flawed economics. Even if most spreadsheet errors are relatively benign, poor spreadsheet practice can risk millions, if not billions, of dollars~\cite{powell2009impact}.

	Fortunately, spreadsheet researchers address these dangers by amassing
	spreadsheet collections from diverse sources, industrial to academic, so
	they can study how people use spreadsheets. Collections like
	EUSES~\cite{fisher2005euses}, Fuse~\cite{barik2015fuse}, and the Enron
	corpus~\cite{hermans2015enron} have already produced valuable research, such
	as the documentation of code smells in spreadsheet
	environments~\cite{hermans2012detecting}~\cite{jansen2015code} and the creation
	of grammars to describe every possible spreadsheet program in
	Excel~\cite{aivaloglou2015grammar}. Additionally, some research compares these
	collections to discover variation in spreadsheet techniques over time and
	location~\cite{jansen2015enron}.

	From this wealth of research and data, we must design tools to better understand
	the datasets and simplify complexities we find in them. One such complexity
	is formula construction; functions are an important part of using spreadsheets
	efficiently, yet the number of potential combinations is massive considering
	that Excel boasts over 400 functions. Tools that explore these combinations in
	practice can aid researchers in understanding how people program and in
	educating new generations of spreadsheets users on how to work best.

	This paper contributes such a tool, \toolnameend, that attempts to fulfill this
	need for better exploration of spreadsheet corpora. Working with spreadsheets
	in Excel, \toolname focuses primarily on the functions, like \texttt{SUM} and \texttt{IF} and
	hundreds more, which are the building blocks of formulae that determine values
	across a spreadsheet. The tool creates an interactive visualization which
	captures how people combine these functions, quantifying the broad patterns of
	function nesting while also capturing anomalous formulae and their locations.

@@

	Other tools focus on creating visualizations external to the sheets. Hermans
	and colleagues address a spreadsheet programmer's information needs by making
	dataflow diagrams from individual sheets through the tools
	GyroSAT~\cite{hermans2011supporting} and Breviz~\cite{hermans2011breviz}.
	\toolname is among this type of tool, creating standalone visualizations about
	spreadsheets instead of drawing connections inside the spreadsheet itself. In
	contrast to the other visualization work, our work focuses on representing
	functions and how they combine into complex formulae, rather than exploring the
	data structures of any particular spreadsheet.

@@

	\subsection{Design Goals} \label{goals} In visualizing the spreadsheet data, we
	set out a few core goals for what the tool should accomplish. With
	these goals, we also considered the likely challenges of achieving them
	and particular outcomes which we considered beyond the scope of the research.
	We present these goals below:
